---
title: "Homework_3"
author: "Dhananjay"
date: "2023-02-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Q1. Get the Feed Grains dataset
  * You will find the dataset here: FeedGrains.csv
  * The first two lines are comments, starting with the '#' character
  * The file has a header and is comma separated
  
```{r}
# Loading the required libraries
library(tidyverse)

# Reading the dataset
feedGrainsData <- read.csv("FeedGrains.csv", header = TRUE, sep = ',', skip = 3)
head(feedGrainsData)
summary(feedGrainsData)

```


## Q2. Get the subset of the data corresponding with only annual AND only seed use data (just one dataset, not two)
  * Use the field SC_Frequency_Desc to find the records associated with annual observations
  * Use the field SC_Attribute_Desc to find the records associated with seed use observations
  * Hint:  the filter() function from dplyr might be useful


```{r}
# Loading the required libraries
library(tidyverse)

# Reading the dataset 
feedGrainsData <- read.csv("FeedGrains.csv", header = TRUE, sep = ',', skip = 3)

# Creating the subset of data corresponding with only annual AND only seed use observations using filter

annualSeedUseSubsettedData <- filter(feedGrainsData, SC_Frequency_Desc == "Annual" & SC_Attribute_Desc == "Seed use" )

head(annualSeedUseSubsettedData)

```


## Q3. Use the subsetted annual seed use data to produce a scatterplot of the year by amount seeded, colored by commodity type
  * Use the field Year_ID for the year and make sure it is on the x-axis
  * Use the field Amount for the amount and it is on the y-axis
  * Use the SC_Commodity_Desc for the commodity type
  * Make sure all plot labels are clear and visible
  
```{r}
# Loading the required libraries
library(tidyverse)
library(ggplot2)

# Reading the dataset 

feedGrainsData <- read.csv("FeedGrains.csv", header = TRUE, sep = ',', skip = 3)

# Creating the subset of data corresponding with only annual AND only seed use observations using filter

annualSeedUseSubsettedData <- filter(feedGrainsData, SC_Frequency_Desc == "Annual" & SC_Attribute_Desc == "Seed use" )

# Plotting the subsetted annual seed use data using the scatterplot of year by amount seeded, colored by commodity type

ggplot(annualSeedUseSubsettedData, aes(x = Year_ID, y = Amount, group = SC_Commodity_Desc)) +
  geom_point(aes(color = SC_Commodity_Desc)) +
  scale_color_manual(values = c("Red", "White", "Green", "Black", "Blue")) +
  labs(title = "Amount Seeded for Various Commodities by Year",
       x = "Year",
       y = "Amount Seeded",
       color = "Commodity Description")


```
  
## Q4. Examine the built-in R dataset, Orange
  * Look at the collection of built-in datasets (hint: ?data)
  * Consider the Orange data set in particular (hint: ?Orange)
  * Describe the mode() and class() of all variables in the data set (hint: ?Orange)

```{r}
# Looking into the collection of built-in R datasets
?data

# List out all the available built-in R datasets 
# data() 

# Looking into the Orange dataset
?Orange
#data(Orange)
print(Orange)

# Describing the mode of all variables in the dataset
mode(Orange$Tree)     
mode(Orange$age)
mode(Orange$circumference)

# Describing the class of all variables in the dataset
class(Orange$Tree)
class(Orange$age)
class(Orange$circumference)
```


## Q5. Plot and analyze the Orange data set
  * Produce one grouped bar plot, where the age and the specific tree number is on the x-axis, while the circumference is on the y-axis -- group by age
  
```{r}
# Loading the required libraries
library(tidyverse)
library(ggplot2)
library(RColorBrewer)

# Converting the Tree and age variable to factor
Orange$Tree <- factor(Orange$Tree, levels = unique(Orange$Tree))
Orange$age <- factor(Orange$age)

# Creating a grouped bar plot with age and tree no. on X-axis and circumference on Y-axis
# Grouped by age

ggplot(Orange, aes(x = Tree, y = circumference, fill = age)) +
  geom_bar(stat = "identity", position = "dodge", color="black") +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Growth of Orange Trees by Age",
       x = "Tree Number",
       y = "Circumference (mm)",
       fill = "Age of the Tree")


```

  * Produce one grouped bar plot, where the age and the specific tree number is on the x-axis, while the circumference is on the y-axis -- group by tree ID number
  
```{r}
# Loading the required libraries
library(tidyverse)
library(ggplot2)

# Converting the Tree and age variable to factor
Orange$Tree <- factor(Orange$Tree, levels = unique(Orange$Tree))
Orange$age <- factor(Orange$age)

# Creating a grouped bar plot with age and tree no. on X-axis and circumference on Y-axis
# Grouped by tree ID number

ggplot(Orange, aes(x = age, y = circumference, fill = Tree)) +
  geom_bar(stat = "identity", position = "dodge", color="black") +
  labs(title = "Growth of Orange Trees by Age",
       x = "Age of the Tree",
       y = "Circumference (mm)",
       fill = "Tree Number") 
  


```


  * Make sure all plot labels are clear and visible, in both plots
  * If you have trouble because something is being treated as a number that you want treated as a factor, recall that you can use the factor() function to convert
  * Suppose you are interested in comparing across ages for a specific tree primarily, and only secondarily across different trees.  Which plot would you prefer?

Answer: The first bar plot is more preferable over the second one, if we are interested in comparing across ages for a specific tree. In the first bar plot, the generated bars are grouped by age which provides users a convenient way to compare the circumference of a specific tree at various ages. 


